---
title: "Replication: The unit effect"
author: "MSM"
date: "2026-09-10"
draft: false
categories: [Replication, Science, Decision Making]
sidebar:
  contents: auto
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA,
                      fig.width = 8,
                      fig.height = 6,
                      fig.align = 'center',
                      fig.path = "static")
library(tidyverse)
library(sjPlot)
library(report)
library(here)

group <- 'BAFS2026'

# Sample description, kept out of the public file
internal <- read_csv(here(paste0('0_data/', group, '_internal.csv')))
# Shareable version, linked below
public <- read_csv(here(paste0('0_data/', group, '_public.csv')))

options(digits = 1)
options(scipen = 999)
```

# Original work

Pandelaere, M., Briers, B., & Lembregts, C. (2011). [How to make a 29% increase look bigger: The unit effect in option comparisons](https://doi.org/10.1086/659000). *Journal of Consumer Research*, 38(2), 308-322.

# Team BA spring 2026

- Bardusha Sarveswaran
- Benno Six
- Celestine Jaloux
- Celine Amherd
- Jorden Lindner
- Karar Al-Ubaidy
- Lena Kowal
- Michelle Zurwerra

# Sample

```{r}
excl_age <-
  internal %>%
  filter(age >= 18 | is.na(age))

excl_finished <-
  internal %>%
  filter(age >= 18 | is.na(age)) %>%
  filter(Finished == TRUE)

excl_duration <-
  internal %>%
  filter(age >= 18 | is.na(age)) %>%
  filter(Finished == TRUE) %>%
  mutate(duration = as.numeric(`Duration (in seconds)`) / 60) %>%
  filter(duration >= 3 & duration <= 12)

bekannt <-
  internal %>%
  select(ID, starts_with('K_'), starts_with('Feedback')) %>%
  filter(Feedback_Q1 == 'Nein')

Kontrollfragen <-
  internal %>%
  select(ID, starts_with('K_'), starts_with('Feedback')) %>%
  filter(Feedback_Q1 == 'Nein') %>%
  mutate(
    check_1 = case_when(K_1 == 'Blau' ~ 1, TRUE ~ 0),
    check_2 = case_when(K_2 == 'fliegen' ~ 1, TRUE ~ 0),
    check_3 = case_when(K_3 == 2 ~ 1, TRUE ~ 0)) %>%
  mutate(sum = (check_1 + check_2 + check_3)) %>%
  group_by(ID) %>%
  reframe(sum = sum(sum)) %>%
  filter(sum > 1)

# Same chain as the Overview: age, completion and duration are applied
# before the attention check decides who stays in.
included <-
  internal %>%
  filter(age >= 18 | is.na(age)) %>%
  filter(Finished == TRUE) %>%
  mutate(duration = as.numeric(`Duration (in seconds)`) / 60) %>%
  filter(duration >= 3 & duration <= 12) %>%
  select(ID, starts_with('K_'), starts_with('Feedback')) %>%
  filter(Feedback_Q1 == 'Nein') %>%
  mutate(
    check_1 = case_when(K_1 == 'Blau' ~ 1, TRUE ~ 0),
    check_2 = case_when(K_2 == 'fliegen' ~ 1, TRUE ~ 0),
    check_3 = case_when(K_3 == 2 ~ 1, TRUE ~ 0)) %>%
  mutate(sum = (check_1 + check_2 + check_3)) %>%
  group_by(ID) %>%
  reframe(sum = sum(sum)) %>%
  filter(sum > 1) %>%
  distinct(ID) %>%
  pull(ID)
```

A total of `r dim(internal)[1]` people took part in the study. The following filters were applied:

1. Under 18 years of age: `r dim(internal)[1]-dim(excl_age)[1]` participants removed
2. Study not completed: `r dim(internal)[1]-dim(excl_finished)[1]` removed
3. Study completed too quickly or too slowly: `r dim(internal)[1]-dim(excl_duration)[1]` removed
4. Already knew the study (answered 'yes'):`r dim(internal)[1]-dim(bekannt)[1]` removed
5. At least 2 of 3 control questions correct: `r dim(internal)[1]-dim(Kontrollfragen)[1]` removed

That leaves `r length(included)` in the sample. Note: the criteria are listed as applied independently, not cumulatively.

```{r}
demo <-
  internal %>%
  filter(ID %in% included) %>%
  select(ID, gender, year, age, job, starts_with('Manipulation'),
         starts_with('Interesse'), starts_with('Feedback'))
```

Of these, `r dim(demo %>% filter(gender == 'Weiblich'))[1]` are female (*M* = `r demo %>% filter(gender == 'Weiblich') %>% summarise(age = round(mean(age, na.rm = TRUE),1))`, *SD* = `r demo %>% filter(gender == 'Weiblich') %>% summarise(age = round(sd(age, na.rm = TRUE),1))` years), `r dim(demo %>% filter(gender == 'Männlich'))[1]` male (*M* = `r demo %>% filter(gender == 'Männlich') %>% summarise(age = round(mean(age, na.rm = TRUE),1))`, *SD* = `r demo %>% filter(gender == 'Männlich') %>% summarise(age = round(sd(age, na.rm = TRUE),1))` years), `r dim(demo %>% filter(gender == 'Divers'))[1]` non-binary (*M* = `r demo %>% filter(gender == 'Divers') %>% summarise(age = round(mean(age, na.rm = TRUE),1))`, *SD* = `r demo %>% filter(gender == 'Divers') %>% summarise(age = round(sd(age, na.rm = TRUE),1))` years) and `r dim(demo %>% filter(is.na(gender)))[1]` did not report a gender.

# Replication Study S3

The same choices presented in kcal versus kJ. Eight tasks each, averaged, with kJ converted to kcal.

```{r fig-s3}
S3 <-
  public %>%
  filter(ID %in% included) %>%
  select(ID, starts_with('S3_kJ_Q'), starts_with('S3_kcal_Q')) %>%
  pivot_longer(
    cols = -ID,
    names_to = 'item',
    values_to = 'answers') %>%
  separate(item, into = c("study", "unit", "item"), sep = "_") %>%
  mutate(choice_value = as.numeric(str_extract(answers, "[\\d']+") %>%
                                     str_remove_all("'"))) %>%
  mutate(choice_value = case_when(
    unit == 'kcal' ~ choice_value,
    unit == 'kJ' ~ choice_value / 4.184,
    TRUE ~ NA_real_)) %>%
  filter(!is.na(answers))

S3 %>%
  group_by(unit) %>%
  summarise(
    mean_value = mean(choice_value, na.rm = TRUE),
    se = sd(choice_value, na.rm = TRUE) / sqrt(n()),
    ci_lo = mean_value - 1.96 * se,
    ci_hi = mean_value + 1.96 * se
  ) %>%
  ggplot(aes(x = unit, y = mean_value, color = unit)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.2) +
  scale_color_viridis_d() +
  labs(
    title = "Average chosen value (95% CI)",
    x = "Unit", y = "kcal"
  ) +
  theme_minimal(base_size = 20) +
  theme(legend.position = "none")
```

A t-test between the two conditions (kcal v. kJ) gives:

```{r}
t.test(choice_value ~ unit, data = S3) |> report()
```

# Replication Study S4

Delivery time presented in days versus months. Ten tasks each, with months converted to days.

```{r fig-s4}
S4 <-
  public %>%
  filter(ID %in% included) %>%
  select(ID, starts_with('S4_ET'), starts_with('S4_EM'),
         starts_with('S4_AET'), starts_with('S4_AEM')) %>%
  pivot_longer(
    cols = -ID,
    names_to = 'item',
    values_to = 'answers'
  ) %>%
  separate(item, into = c("study", "unit", "item"), sep = "_") %>%
  na.omit() %>%
  mutate(choice_delay = as.numeric(str_extract(answers, "\\d+"))) %>%
  mutate(days = case_when(
    unit == 'EM' | unit == 'AEM' ~ choice_delay * 31,
    TRUE ~ choice_delay
  ))

S4 %>%
  group_by(unit) %>%
  summarise(
    mean_days = mean(days, na.rm = TRUE),
    se = sd(days, na.rm = TRUE) / sqrt(n()),
    ci_lo = mean_days - 1.96 * se,
    ci_hi = mean_days + 1.96 * se
  ) %>%
  ggplot(aes(x = unit, y = mean_days, color = unit)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.2) +
  labs(
    title = "Average days per condition (95% CI)",
    x = "Unit", y = "Days"
  ) +
  scale_color_viridis_d() +
  theme_minimal(base_size = 20) +
  theme(legend.position = "none")
```

A t-test between the two conditions (days v. months) gives:

```{r}
S4_TM <-
  S4 %>%
  filter(unit == 'ET' | unit == 'EM')

t.test(days ~ unit, data = S4_TM) |> report()
```

A t-test between the two conditions (months v. attention months) gives:

```{r}
S4_AET <-
  S4 %>%
  filter(unit == 'EM' | unit == 'AEM')

t.test(days ~ unit, data = S4_AET) |> report()
```

# Data

The shared version of the dataset is available as [`BAFS2026_public.csv`](../../0_data/BAFS2026_public.csv). It contains the experimental responses without gender, year of birth and age. For analyses that need those fields, please contact the author.
